|

### 2022-01-26 text/ManuscriptDraft.Rmd
Manuscript update

If interacting and interdependent species, e.g. plants and pollinators, react to different environmental cues, or react differently to the same cues, their phenology may be desynchronized and temporal mismatches occur. Changes in flower phenology as a consequence of climate change may indirectly affect the abundance of flower-dependent species [@hoye2013] and these effects can be stronger than the direct effects of climate change [@ogilvie2017]. However, the relation between flower phenology and reproductive success contain many facets and is not simply a question of temporal overlap with pollinators. For example, @bolmgren2015 found that for an insect-pollinated plant species early flowering was associated with increased fruit-set but late flowering was associated with higher seed mass. Further, periods of resource scarcity through the season may be more important than the floral sum of the season, i.e. dynamics in resource availability above a level of sufficiency may have a weak effect on pollinator abundance while the time above this threshold has a stronger effect [@ogilvie2017].

Different phenological factors, such as onset, peak, and end of flowering may respond differently to climate change [kilder fra første manuskript] and estimation of the degree of mismatch between species may thus be sensitive to the choice of metric [@inouye2019]. Efficient and precise methods for deriving full seasonal distributions of phenology at the level of individuals is much needed but usually prevented for logistical reasons.

@@

#
## The image series and flower annotations
Images were collected throughout the full growing season. For this study, we limited the 1-hr interval image-series to the flowering season and randomly randomly sampled images from each series. The number of images for each series after sampling is given in table 1. All flowers in these images were manually annotated using the rectangular bounding box tool in the VIA VGG annotation software [@dutta2019] and each individual flower was assigned a unique ID. These annotations constitute our ground truth tracks. The series varied substantially in the number of flowers they contained (table 1). Figure 1 show annotated example images from three series (xxx) at different times during the flowering season (TO-DO).

@@

Several issues complicate the task of tracking individual flowers through time-lapse image-series. First, stand-alone monitoring of the relatively slow process of flower development through full growing seasons (several months) requires low frame rates (several images per day). These frame rates capture the phenology of the flowers at very high detail, but fail to capture the full movement of flowers. For example, as the wind shifts, the flower heads changes direction, which in the image series happens instantaneously (i.e., between two consecutive frames). However, as they are constrained by their stalk, there is a limit to the distance they can move. With the tracking parameter **running mean**, we base the tracking on the distance between a point in the current frame and the running mean of the positions of a number of previous points in a track, taking advantage of the fact that flower move around a central point.

Second, as they move, flowers may temporarily occlude the view of other flowers and flowers close to the edge of the frame may move in and out of view. This can cause a track to be lost and a new track erroneously being initiated when the flower reappears. However, if a flower reappears in the same area as a flower is already being tracked after disappearing in a few frames, it is reasonable to assume that it is the same individual and not that the old flower wilted/disappeared and a new one developed. The parameter **max gap size** sets the number of frames a track can be lost before a new track is initiated for points appearing in the same area. Concurrently, this deals with potential false negatives. If a given flower has not been detected in a few frames, the track will not be lost. The counter for number of disappeared frames for a track is reset when a new point is associated with the track within the threshold.

Third, a centroid tracking algorithm will associate a point with a tracking based on minimum distance only, disregarding the absolute distance. As flowers are constrained in the movement, we can assume a maximum distance that will occur between points and force the initiation of tracks for points that exceed this threshold. We do that with the parameter **max distance**.

@@

To explore the effect of the user parameters and to identify the optimal combination of parameters for our case of tracking flowers, we applied the tracking algorithm on each of the five image-series with every combination of a range of values for each parameter (3179 combinations): Values for max gap size and running mean were 0-160 with a step size of 10 and for max distance we used values 0-1000 with a step size of 100. Note that max distance set to zero ignores the parameter altogether. We identified the setting(s) that returned the lowest number of track mismatches and compare the tracking results between optimal settings and all parameters set to zero.

@@

The commonly used multiple object tracking accuracy (MOTA) score quantifies tracking performance based on counts of tracking mismatches [@bernardin2008]. Mismatches occur when objects swap track identity because they are in close vicinity to each other or when an object periodically disappears and is assigned a new track identity when it reappears. Only the shifts in tracking identity are counted as mismatches while the number of points assigned to each track is not considered. We use three evaluation metrics: We calculate the number of mismatches and derive the MOTA score (number of mismatches in relation to the total number of points) to evaluate performance of our tracking and filtering algorithms. Finally, we compare the number of tracks identified by the automatic tracking with the true number of flowers in a series. These should ideally be equal. However, a correct number of tracks (i.e. number of tracks equal to number of objects) does not necessarily translate to correct tracking, since individual tracks can be erroneously split by the tracking algorithm and several tracks can be combined. We present the tracking performance by both the lowets number of mismatches/highest MOTA given overall and for combinations that returned a number opf tracks equal to the number of flowers in the series.

We base our filtering on a density-based clustering of the centroid of a geometry of each track. The geometries are derived in the following way: For tracks consisting of only a single point, the coordinates of the point are used as the centroid. For tracks consisting of two points, we establish the straight line between the points and calculate the centroid of the line. For tracks consisting of three points, we establish the triangle from the points and calculate the centroid of the triangle. For tracks consisting of three points or more, we calculate the convex hull of all the points included in the track and derive the polygon from the vertices of the convex hull and calculate the centroid of this polygon. For tracks that contain more than two points but where the points a colinear, we establish the line through the points and calculate the centroid of the line.

We then apply the filtering based on the geometry centroids to remove tracks in areas with a high density of tracks as these have a high risk of tracking mismatches. The filtering is done in two steps. First, the DBSCAN clustering algorithm [@ester1996] is run on the centroids with a conservatively high value for the eps parameter (350), meaning that tracks in close vicinity to each other will be clustered together. The value for eps was chosen by applying the filtering algorithm across the three series at range of eps values from x to x and a step size of x and identifying the value that returned the highest number of tracks across the three series at the lowest number of mismatches.

Second, all tracks that were not assigned to a unique cluster are removed. Tracks that are spatially isolated remains. We evaluate the tracking accuracy of the remaining tracks. We demonstrate our filtering approach for the image series for which our tracking algorithm did not produce perfect results.

@@

The results of the tracking parameter test are given in table 1. We present the tracking performance without the use of the three tracking parameters along with the best performance (lowest number of mismatches/highest MOTA), and the best performance where the number of tracks correspond to the number of individual flowers in the series.

**Table 1:** Table description...
#
![](../figures/table1.png){ width=100% }

@@

The results of the filtering algorithm on the three series in which our tracking algorithm did not return perfect results are given in table 2 and the filtering process is visualised in fig. 4. Our filtering method successfully extracted 28 tracked flowers with only a single mismatch from the three series with complex scenes using a fixed value of 350 for eps and fixed values for the tracking parameters (10, 10, 300, for running mean, max gap size and max distance, respectively).

**Table 2:** Results of the filtering algorithm with an eps value of 350 applied across all three . `r word_comment("Note to self. Table shows to what degree a tr track contains points from a single gt track, but does not show whether all points of the gt track are included in the tr track. We know that is the case, but should it be tested/shown?", "Table 1")` series.

@@

Many tracking algorithms, such as the Simple online and real-time tracking (SORT) [@sort], use Kalman filtering to predict the future location of an object by estimating its velocity [@kalmann]. However, Kalman filtering assumes linear motion of the object and does not handle abrupt motion well. As the flowers are constrained in movement by their stalk, their movement does not follow the assumption of linear motion. Some tracking algorithms, such as DeepSORT [@deepsort2017], take the tracking-by-detection approach in which convolutional neural networks (CNNs) are applied to leverage appearance information in order to continuously distinguish individuals through time, thus aiding the tracking algorithm. In our case, however, flowers are not clearly distinguishable and one flower may look less alike between two frames than two flowers in a single frame.

Fine-tuning and evaluating any tracking algorithm requires a ground dataset to compare tracking accuracy against.

We applied our tracking algorithm on 3179 combinations of the three parameters. Using the parameters increased the performance of the tracking substantially. Our steps in parameter values were crude, however, and it is very likely that finer steps in these values would identify combinations that produce even better results. Here we do not perform this analysis, however, as our goal is to show that the parameters can be used for optimizing tracking performance in general.

In a naive setting, a priori knowledge about which settings to use is not available.

As with any tracking method, accuracy can only be assessed based on visual ... or more ... based on a subsample of tracks for whih a ground truth exists.

We recommend to perform a parameter test as demonstrated here on subset of ground truth tracks to derive optimal parameters for application of the method in a new setting.

Depending on the nature of the objects being tracked and the complexity of the scene, the user parameters can be estimated from visual examination of the tracking results. Often it may be preferable to manually annotate a subset of the objects in the image series and derive a set of user parameters from these results.

Number of tracks equal to number of flowers does not necessarily mean that the tracking is correct. I.e., the tracking algorithm could split one ground truth track in two, but combine two other ground truth tracks into one, which would even each other out.

Our tracking algorithm consistently returns high MOTA scores.

For complex scenes, all three parameters make a difference.

The parameters are interdependent.

In cases where we are tracking perfectly with maxDisap = 0, setting it any value will not make a difference. Not quite right. Explore more...

@@

Extracting tracks that are spatially isolated does not guarantee that the tracks are correct/without errors. However, as spatially isolated objects are easier to track, it increases confidence in the remaining tracks.

**Alekos: I think it would be good to add a short description or comment on what each of the mentioned parameters is used for (e.g. why did you need to change its value, how this affects the performance of the method on the challenges in your dataset compared to standard computer vision tracking datasets).**

We applied a single value for the DBSCAN eps parameter in our tracking algorithm. We note that this value could be fine-tuned for improved results for the individual series (i.e. more flowers extracted without increased number of mismatches). However, as our goal here was to show that a single conservative value can be applied across series, we do not show those results here. In a naive setting, a general value could be chosen or the value could be adjusted for each series based on visual examination of performance or testing on a subset of data.

Our method for filtering tracks using DBSCAN on track centroids ensures that all tracks are given the same weighting in the filtering since each track is represented by a single point. In some cases 
e.g. if it is given a priori that an object will always appear in a minimum of two frames, then single point tracks can be filtered out. However, when such a priori knowledge is not accessible, a conservative approach as the one we present is preferable.

>>>>

If interacting and interdependent species, e.g. plants and pollinators, react to different environmental cues, or react differently to the same cues, their phenology may be desynchronized and temporal mismatches occur. Changes in flower phenology as a consequence of climate change may indirectly affect the abundance of flower-dependent species [@hoye2013] and these effects can be stronger than the direct effects of climate change [@ogilvie2017]. However, the relation between flower phenology and reproductive success is multifaceted and not simply a question of temporal overlap with pollinators. For example, @bolmgren2015 found that for an insect-pollinated plant species early flowering was associated with increased fruit-set but late flowering was associated with higher seed mass. Further, periods of resource scarcity through the season may be more important than the floral sum of the season, i.e. dynamics in resource availability above a level of sufficiency may have a weak effect on pollinator abundance while the time above this threshold has a stronger effect [@ogilvie2017].

Different phenological factors, such as onset, peak, and end of flowering may respond differently to climate change [@cook2012] and estimation of the degree of mismatch between species may thus be sensitive to the choice of metric [@inouye2019]. Efficient and precise methods for deriving full seasonal distributions of phenology at the level of individuals is much needed but so far most often unfeasible for logistical reasons.

@@

#
## The image series and flower annotations
Images were collected throughout the full growing season. While the original time-lapse frequency was higher, the basis for this study was 1-hr interval sereis. We limited the 1-hr interval image-series to the flowering season and randomly randomly sampled images from each series. The number of images for each series after sampling is given in table 1. All flowers in each of these images were manually annotated using the rectangular bounding box tool in the VIA VGG annotation software [@dutta2019] and each individual flower was assigned a unique ID. These annotations constitute our ground truth tracks. The series varied substantially in the number of flowers they contained (table 1). Figure 1 show annotated example images from three series (xxx) at different times during the flowering season (TO-DO).

@@

Several issues complicate the task of tracking individual flowers through time-lapse image-series.

First, stand-alone monitoring of the relatively slow process of flower development through full growing seasons (several months) requires only low frame rates. Here we used several images per day. These frame rates capture the phenology of the flowers at very high detail, but fail to capture the full movement of flowers. For example, as the wind shifts, the flower heads change direction, which in the image series happens instantaneously (i.e. between two consecutive frames). However, as they are constrained by their stalk, there is a limit to the distance they can move. With the tracking parameter **running mean**, we have the option to base the tracking on the distance between a point in the current frame and the running mean of the positions of a number of previous points in a track, taking advantage of the fact that flower move around a central point.

Second, as flowers move, they may temporarily occlude the view of other flowers and flowers close to the edge of the frame may move in and out of view. This can cause a track to be lost and a new track erroneously being initiated when the flower reappears. However, if a flower reappears in the same area as a flower is already being tracked after disappearing in a few frames, it is reasonable to assume that it is the same individual and not that the old flower wilted/disappeared and a new one developed. The parameter **max gap size** sets the number of frames a track can be lost before a new track is initiated for points appearing in the same area. If a given flower has not been detected in a few frames, the track will not be lost. Concurrently, this deals with potential false negatives. The counter for number of disappeared frames for a track is reset when a new point is associated with the track within the threshold.

Third, a centroid tracking algorithm will associate a point with a tracking based on minimum distance only, disregarding the absolute distance. As flowers are constrained in the movement, we can assume a maximum distance that will occur between points and force the initiation of tracks for points that exceed this threshold. We set this threshold with the parameter **max distance**.

@@

To explore the effect of the user parameters and to identify the optimal combination of parameters for our case of tracking flowers, we applied the tracking algorithm on each of the five image-series with every combination of a range of values for each parameter (3179 combinations): Values for max gap size and running mean were 0-160 with a step size of 10 and for max distance 0-1000 with a step size of 100. Note that max distance set to zero ignores the parameter altogether. We identified the setting(s) that returned the lowest number of track mismatches and compare the tracking results between optimal settings and all parameters set to zero.

We base our filtering on a density-based clustering of the centroid of a geometry of each track. The geometries are derived in the following way: For tracks consisting of only a single point, the coordinates of the point are used as the centroid. For tracks consisting of two points, we establish the straight line between the points and calculate the centroid of the line. For tracks consisting of three points, we establish the triangle from the points and calculate the centroid of the triangle. For tracks consisting of four points or more, we calculate the convex hull of all the points included in the track and derive the polygon from the vertices of the convex hull. Hereafter we calculate the centroid of the polygon. For tracks that contain more than two points but where the points a collinear, we establish the line through the points and calculate the centroid of the line.

The filtering is done on the geometry centroids to remove tracks in areas with a high density of tracks as these have a high risk of tracking mismatches. The filtering is done in two steps. First, the DBSCAN clustering algorithm [@ester1996] is run on the centroids with a conservatively high value for the eps parameter, meaning that tracks in close vicinity to each other will be clustered together. Second, all tracks that were not assigned to a unique cluster are removed. Tracks that are spatially isolated remains.

We demonstrate the filtering method on the three series for which our tracking algorithm did not return perfect results. For this, we chose a fixed combination of the tracking algorithm parameters across all three series (max gap size: 10; running mean: 10; max distance: 300).

We then applied the filtering method on the three series with a range of values for eps (x-xx, step size of 10) and identified the value that returned the highest sum of tracks across the three series at the lowest number of mismatches. We present the tracking accuracy before filtering and on the remaining tracks after filtering. We underline that the results presented are thus highly conservative. I.e., the approach could be optimised to extract a higher number of tracks while maintaining the number of mismatches. For example, basing the filtering on the best performing tracking parameters for each series specifically would improve the output of the tracking algorithm. Similarly, the results could be improved by fine-tuning the value for eps for each series individually. However, as our goal here was to demonstrate that our tracking and filtering method can be applied in a naive setting with fixed values for each parameter we have not included those results.

@@

The commonly used multiple object tracking accuracy (MOTA) score quantifies tracking performance based on counts of tracking mismatches [@bernardin2008]. Mismatches occur when objects swap track identity because they are in close vicinity to each other or when an object periodically disappears and is assigned a new track identity when it reappears. Only the shifts in tracking identity are counted as mismatches while the number of points assigned to each track is not considered. We use three evaluation metrics: We calculate the number of mismatches and derive the MOTA score (number of mismatches in relation to the total number of points) to evaluate performance of our tracking and filtering algorithms. Finally, we compare the number of tracks identified by the automatic tracking with the true number of flowers in a series. These should ideally be equal. However, a correct number of tracks (i.e. number of tracks equal to number of objects) does not necessarily translate to correct tracking, since individual tracks can be erroneously split by the tracking algorithm and several tracks can be combined. We present the tracking performance by both the lowest number of mismatches/highest MOTA given overall and for combinations that returned a number of tracks equal to the number of flowers in the series.

@@

The results of the tracking parameter test are given in table 1. We present the tracking performance without the use of the three tracking parameters along with the best performance (lowest number of mismatches/highest MOTA), and the best performance where the number of tracks correspond to the number of individual flowers in the series. When several values for a parameter was included in a result, we present the range of values. For two series, 2018 THUL B and 2019 NARS D, the tracking algorithm returned perfect results for several combinations of parameters. For the three remaining series, the algorithm returned consistently high MOTA scores (above 0.95). In all cases, except 2019 NARS D, where the tracking algorithm returned perfect results with all parameters zeroed, did the inclusion of the user set tracking parameters improve the tracking accuracy.

In one case, NARS-13, four different numbers of tracks were returned among the parameter combinations that resulted in the lowest number of mismatches. A number of tracks equal to the number of flowers in the series were among these four.

In one case, NYAA-04, the parameter combination that produced the highest accuracy did not return a number of tracks equal to the number of flowers in the series. However, combinations that did produce this number of tracks also returned a higher number of mismatches.

**Table 1:** Table description... To-do: Rename series to camera codes.

#
![](../figures/table1.png){ width=100% }

@@

The results of the filtering algorithm on the three series in which our tracking algorithm did not return perfect results are given in table 2 and the filtering process is visualised in fig. 4. Our filtering method extracted 28 tracked flowers from the three series with complex scenes using a fixed value of 350 for eps and fixed values for the tracking parameters (10, 10, 300, for running mean, max gap size, and max distance, respectively) with only a single mismatch included in the 28 tracks.

**Table 2:** Results of the filtering algorithm with an eps value of 350 applied across all three series To-do: Change image series to camera codes.

@@

Many tracking algorithms, such as the Simple online and real-time tracking (SORT) [@sort], use Kalman filtering to predict the future location of an object by estimating its velocity [@kalmann]. However, Kalman filtering assumes linear motion of the object and does not handle abrupt motion well. As the flowers are constrained in movement by their stalk, their movement does not follow the assumption of linear motion. Some tracking algorithms, such as DeepSORT [@deepsort2017], take the tracking-by-detection approach in which convolutional neural networks are applied to leverage appearance information in order to continuously distinguish individuals through time, thus aiding the tracking algorithm. In our case, however, flowers are not clearly distinguishable and one flower may look less alike between two frames than two flowers in a single frame.

Finally, some tracking methods require manual initiation of a fixed number of tracks for objects in the first frame and the algorithm will track these objects in the subsequent frames [@hu2012]. These methods does not allow new objects to enter the frame through the series. In our case, objects can disappear permanently from the frame, for example when flowers wilt, in which case the track should be terminated. Oppositely, new objects can continuously appear through the series when new flowers bloom of previously unseen flowers enter the frame. Additionally, objects may temporarily disappear from view and may or may not reappear at some later stage. These settings vastly complicate the task of tracking multiple objects.

Our tracking algorithm returned consistently accuracy despite the difficult settings.

We applied our tracking algorithm on 3179 combinations of the three parameters. Using the parameters increased the performance of the tracking substantially. Our steps in parameter values were crude, however, and it is very likely that finer steps in these values would identify combinations that produce even better results. Here we do not perform this analysis, however, as our goal is to show that the parameters can be used for optimizing tracking performance in general

In a naive setting, a priori knowledge about which settings to use is not available. As with any tracking method, accuracy can only be assessed based on visual observation or more thoroughly by testing on subsample of tracks for which a ground truth exists. When applying the method in a new setting, we recommend performing a parameter test as demonstrated here on a subset of ground truth tracks to derive the optimal set of parameters.

Number of tracks equal to number of flowers does not necessarily mean that the tracking is correct. I.e., the tracking algorithm could split one ground truth track in two, but combine two other ground truth tracks into one, which would even each other out. This was the case for NYAA-04 for which the parameter combination that returned a number of tracks equal to the number of flowers in the series was not the combination the produced the highest MOTA score.

The two series for which the tracking algorithm produced perfect results were also the series with the lowest number of flowers. As the number of flowers increase so does the complexity of the task of tracking. A higher number of flowers within the frame increases the possibility of flowers in close vicinity to each other and of flowers overlapping and moving in and out of frame translating to a higher risk of making tracking errors. The very high complexity in the scenes with many flowers, especially NARS-13 with 85 flowers, renders it unlikely that any tracking method could produce perfect results. Although this setting can be considered an extreme case for our application, it underlines the relevance of a pipeline including both tracking and filtering.

We presented a filtering method based on the density of tracks within the frame which can be applied subsequent of tracking. Extracting tracks that are spatially isolated does not guarantee that the tracks are correct/without errors. However, as spatially isolated objects are easier to track, it increases confidence in the remaining tracks.

The filtering algorithm successfully extracted correctly tracked flowers from each of the three series for which the tracking algorithm did not produce perfect results. Only a single mismatch was included in the 28 flowers that passed the filtering.

We applied the filtering algorithm on tracks produced with a fixed set of tracking parameters as well as a fixed value for the DBSCAN eps parameter for the filtering algorithm. We note that the results could be improved (i.e. more flowers extracted without increased number of mismatches) by chosing the best tracking result for each series and fine-tuning the eps value for each individual series. However, as our goal here was to show that a single conservative value can be applied across series for which the tracking algorithm did not necessarily produce the best results, we do not show those results here. As in the case of the tracking parameters, we recommend setting the value for eps based on either visual examination of results or on tests on a subset of ground truth data when applying the method in a new setting.

We note that our method for filtering tracks using DBSCAN on track centroids ensures that all tracks are given the same weighting in the filtering since each track is represented by a single point. In some cases this may be suboptimal. For example, if it is given a priori that an object will always appear in a minimum of two frames, then single point tracks can be filtered out. However, when such a priori knowledge is not accessible, a conservative approach as the one we present is preferable.

## The tracking and filtering parameters

Our method relies on a number of parameters set by the user.

For the three series with complex scenes, applying the max gap size parameter improved tracking accuracy with values of 10-20 being optimal. These series all contain flowers at the edge of the frame and the higher number of flowers increases the risk of flowers temporarily occluding others. Correct tracking of these objects requires that an object can disappear from view without the track instantly being lost, which the max gap size allows. Oppositely, a higher number of flowers also increase the possibility of a new flower blooming in close vicinity to where a flower previously existed and setting the value too high may force objects such objects that are only separated in time but occur in the same space into the same track.

In all cases, except NARS-17, for which perfect tracking was achieved with all parameters set to zero, did a value for running mean above 0 improve results.

Interestingly, setting the value to 10 produced the best results in all these cases, except for the combination returning the lowest number of mismatches for NYAA-04.

Increasing the value for running mean means that less weight is put on movements between each frame.

Maybe NYAA-04 flowers move more? Maybe longer stalk?

Coincidently, the flowers in NYAA-04 also flower longer on average than the other series.

If the average euclidean distance that a flower moves is higher for NYAA-04 it may improve results to even out this movement by increasing the value for running mean.

Look at this tomorrow....

These parameters are not independent. For example, the distance between a point in the current frame and a point in the previous frame may not be the same as that for a point in the current frame and the mean location of the previous ten points in a track. This underlines the relevance of search for an optimal combination based on ground truth data.

Pointe: Selvom vi kigger på mange komvinationer af parameterne, så falder kombinationerne hvor de bedste resultater produceres (laveste mismatches) indenfor forholdsvis smalle områder. Det tyder på at metoden med de givne parametre er generaliserbar og kan anvendes på andre kameraer med samme setting.

For our application, we find that a

For the three series for which the tracking algorithm failed to produce perfect results, the range

Why do they make a difference:

max gap size:

Running mean:

Max distance:

eps:

Offline and online tracking differ

Online tracking

Our tracking algorithm does not rely on information from future frames and can thus be applied both offline and online.

MOT can also be categorized into online tracking and offline
tracking. The difference is whether or not observations from future frames are utilized when handling the current frame. Online, also called causal, tracking methods only rely on the past information available up to the current frame, while

It also depends on the time the flowers are spread over. Many flowers and short time is a hard problem.

@@
